Circumnuclear Media and Accretion Rates of Quiescent Supermassive Black Holes

### 2015-04-09 ms.tex
intro

Knowledge of how $\dot{M}$ depends on the SMBH mass, $\Mbh$, and other
properties of the nucleus informs key questions related to the
co-evolution of SMBHs and their host galaxies with cosmic time
(e.g.~\citealt{Kormendy&Ho13}; \citealt{HeckmanBest:2014a}).  In the
low redshift universe, SMBH growth is dominated by low mass black
holes, $M_{\bullet} \lesssim 10^{8}M_{\odot}$
(e.g.~\citealt{Heckman+04}), a fact often attributed to the trend of
`cosmic down-sizing' resulting from hierarchical structure growth
(e.g, \citealt{Gallo+08}).  However, the physical processes by which typical
low mass black holes accrete could in principle be distinct from those
operating at higher SMBH masses, or those in AGN.  One key question is whether SMBHs grow primarily by the
accretion of gas fed in directly from galactic or extra-galactic
scales, or whether significant growth can result also from local
stellar mass loss in the nuclear region.

@@

In this paper we instead focus on time-independent models, with
sufficient heating to make radiative cooling negligible.  This
approach allows us to systematically explore the relevant parameter
space and to derive analytic expressions that prove useful in
determining whether cooling instabilities manifest across the expected
range of galaxy properties, or whether other (non-AGN) forms of
feedback can produce a prolonged state of steady, thermally stable
accretion.  Even if cooling instabilities develop on galactic scales
over longer $\sim$ Gyr timescales, a quasi steady-state may exist
between these inflow events on smaller radial scales comparable to the
sphere of influence.  We are thus motivated to explore whether such
steady flows are generally expected within the nucleus, or
whether in most galaxies the nuclear gas content is always building
towards instability.
%A chief goal of our analysis is to provide a framework for interpreting the properties of quiescent galactic nuclei given the comparatively well understood physics of stellar winds and supernova heating.

In the presence of strong heating, one-dimensional steady-state flow
is characterized by an inflow-outflow structure, with a critical
radius known as the``stagnation radius" $\rs$ where the radial
velocity passes through zero.  Mass loss from stars interior to the
stagnation radius is accreted, while that outside $\rs$ is unbound in
an outflow from the nucleus.  The stagnation radius, rather than the
Bondi radius, thus controls the SMBH accretion rate (although we will
show that $\rs$ will in many cases reside near the nominal Bondi
radius).  When heating is sufficiently weak, however, the stagnation
radius may move to much larger radii or not exist at all, in which
case nearly the entire ISM of the galaxy is flowing towards the SMBH.
A goal of this work is to assess under what conditions a stagnation
radius exists, as this question is intimately tied to that of cooling
instability.

>>>>

Knowledge of how $\dot{M}$ depends on the SMBH mass, $\Mbh$, and other
properties of the nucleus informs key questions related to the
co-evolution of SMBHs and their host galaxies with cosmic time
(e.g.~\citealt{Kormendy&Ho13}; \citealt{HeckmanBest:2014a}).  In the
low redshift universe, SMBH growth is dominated by low mass black
holes, $M_{\bullet} \lesssim 10^{8}M_{\odot}$
(e.g.~\citealt{Heckman+04}), a fact often attributed to the trend of
`cosmic down-sizing' resulting from hierarchical structure growth
(e.g, \citealt{Gallo+08}).  However, the physical processes by which typical
low mass black holes accrete could in principle be distinct from those
operating at higher SMBH masses, or those in AGN.  Of key importance is whether SMBHs grow primarily by the
accretion of gas fed in directly from galactic or extra-galactic
scales, or whether significant growth can result also from local
stellar mass loss in the nuclear region.

@@

In this paper we instead focus on time-independent models, in cases when the nuclear gas receives sufficient heating to render radiative cooling negligible.  This approach allows us to systematically explore the relevant parameter
space and to derive analytic expressions that prove useful in
determining whether cooling instabilities manifest across the expected
range of galaxy properties, or whether other (non-AGN) forms of
feedback can produce a prolonged state of steady, thermally stable
accretion.  Even if cooling instabilities develop on galactic scales
over longer $\sim$ Gyr timescales, a quasi steady-state may exist
between these inflow events on smaller radial scales comparable to the
sphere of influence.  We are thus motivated to explore whether such
steady flows are generally expected within the nucleus, or
whether in most galaxies the nuclear gas content is always building
towards instability.
%A chief goal of our analysis is to provide a framework for interpreting the properties of quiescent galactic nuclei given the comparatively well understood physics of stellar winds and supernova heating.

In the presence of strong heating, one-dimensional steady-state flow
is characterized by an inflow-outflow structure, with a critical
radius known as the``stagnation radius" $\rs$ where the radial
velocity passes through zero.  Mass loss from stars interior to the
stagnation radius is accreted, while that outside $\rs$ is unbound in
an outflow from the nucleus.  The stagnation radius, rather than the
Bondi radius, thus controls the SMBH accretion rate (although we will
show that $\rs$ will usually reside near the nominal Bondi
radius).  When heating is sufficiently weak, however, the stagnation
radius may move to much larger radii or not exist at all, in which
case nearly the entire ISM of the galaxy is flowing towards the SMBH.
Here we assess under what conditions a stagnation
radius exists, a question intimately tied to that of cooling
instability.
